---
title: "Data Science Tools and Techniques"
author: "Veronika Cintulova"
date: "2023-01-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries used:

```{r}
library('tidyverse')
library('dplyr')
library('psych')
library('purrr')
library('mice')
library('texreg')
library('ggplot2')
```

## Data preparation

Making a copy out of the original dataset and deleting missing values

```{r}
DF <- Life_Expectancy_Data
DFs <- DF
DFs <- na.omit(DFs)
```

Creating a new, joined dataset, which consist out of the original dataset and additional dataset with missing values

```{r}
joined <- merge(DF, World_bank, by= c("Country", "Year"),all.x = TRUE, all.y = TRUE)
DF1 <- unite(joined,"Population", Population.x,Population.y, na.rm = TRUE, remove = TRUE )
DF1 <- unite(DF1, "GDP", GDP.x, GDP.y, na.rm = TRUE, remove= TRUE)
DF2 <- DF1
DF2 <- na.omit(DF2)
```

Creating a new dataset where the character values are deleted = it is easier for models, correlation matrix and they are not needed

```{r}
DF3 <- select(DF2, -"Country", -"Status")
DF3 <- lapply(DF3,as.numeric)
DF4 <- na.omit (DF3)
```

Creating a correlation matrix

```{r}
describe(DF2)  
DF4 <- char2numeric(DF2)  
fa(DF4, nfactors=3) 
cor(DF4)
```

### Data preparation for Question 3

Creating a new dataset with life expectancy lower than 65

```{r}
Dfm <- DF4
Dfv <- Dfm %>% filter(Life_expectancy <65)
```

Deleting char. variables (county, year), and also year, population and GDP as they are not relevant

```{r}
Dfv <- select(Dfv, - c(1:3, 17:18))
```

```{r}
cor(Dfv)
```

### Data preparation for Question 8

Creating a new dataset with only densely populated countries

```{r}
Dfc <- DF2
Dfk <- filter(Dfc, Country %in%  c("Monaco", "Singapore","Hong Kong","Gibraltar","Bahrain","Maldives","Malta","Bermuda","Bangladesh", "China"))
```

Creating a new dataset without the densely populated countries

```{r}
DFa <-DF2
filter(DFa, !Country %in% c("Monaco", "Singapore","Hong Kong","Gibraltar","Bahrain","Maldives","Malta","Bermuda","Bangladesh", "China"))
```

## Modeling

### Question 1 and 2

Creating the first two models to see which variables and how affect the life expectancy.

```{r}

model1Posit <-lm(scale(Life_expectancy)~scale(Alcohol)+scale(percentage_expenditure)+scale(Hepatitis_B)+scale(BMI)+scale(Polio)+scale(Total_expenditure)+scale(Diphtheria)+scale(Population)+scale(Income_composition_of_resources)+scale(Schooling), data=DF3)
summary(model1Posit)

model1Negat <-lm(scale(Life_expectancy)~scale(Adult_Mortality)+scale(infant_deaths)+scale(Measles)+scale(under.five_deaths)+scale(HIV.AIDS)+scale(GDP)+scale(thinness._1.19_years), data=DF3)
summary(model1Negat)
```

Creating a table out of the first model

```{r}
screenreg(list(model1Posit, model1Negat),
          custom.model.names = c("Values with Positive, and ","Negative correlation"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))
```

### Question 3

Creating a model to see how immunization would help Life expectancy when lower than 65

```{r}
model3 <- lm(scale(Life_expectancy)~scale(Hepatitis_B)+scale(Diphtheria)+scale(Polio), data = Dfv)
summary(model3)
screenreg(list(model3),
          custom.model.names = c("Model 3"),
          custom.coef.names = c("Intercept", "Hepatitis_B","Diphtheria","Polio"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))
```

Creating three models (H,D,P) to see how immunization would help Life expectancy when lower than 65

```{r}
modelH <- lm(scale(Life_expectancy)~scale(Hepatitis_B), data = Dfv)
summary(modelH)
screenreg(list(modelH),
          custom.model.names = c("Hepatitis B"),
          custom.coef.names = c("Intercept", "Hepatitis_B"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))

modelD <- lm(scale(Life_expectancy)~+scale(Diphtheria), data = Dfv)
summary(modelD)
screenreg(list(modelD),
          custom.model.names = c("Diphtheria"),
          custom.coef.names = c("Intercept", "Diphtheria"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))

modelP <- lm(scale(Life_expectancy)~+scale(Polio), data = Dfv)
summary(modelP)
screenreg(list(modelP),
          custom.model.names = c("Polio"),
          custom.coef.names = c("Intercept", "Polio"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))
```

```{r}
ggplot(Dfv, aes(x = Hepatitis_B, y = Life_expectancy)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red")

ggplot(Dfv, aes(x = Diphtheria, y = Life_expectancy)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red")

ggplot(Dfv, aes(x = Polio, y = Life_expectancy)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red")
```

Creating a model to see how diseases and other health related variables are affecting the Life expectancy when lower than 65

```{r}
model4 <- lm(scale(Life_expectancy)~scale(Measles)+scale(BMI)+scale(HIV.AIDS), data = Dfv)
summary(model4)
screenreg(list(model4),
          custom.model.names = c("Diseases"),
          custom.coef.names = c("Intercept", "Measles","BMI","HIV.AIDS"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))
```

### Question 4

Creating a model to see how adult mortality affects the life expectancy

```{r}
model5 <- lm(scale(Life_expectancy)~scale(Adult_Mortality), data = DF3)
summary(model5)
screenreg(list(model5),
          custom.model.names = c("Adult Mortality"),
          custom.coef.names = c("Intercept", "Adult_Mortality"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))
```

Ploting a scatterplot with a regression line

```{r}

ggplot(DF2, aes(x = Adult_Mortality, y = Life_expectancy)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red")
```

### Question 5

Creating a model to see how alcohol and BMI (bad lifestyle) affect the life expectancy

```{r}
model6 <- lm(scale(Life_expectancy)~scale(Alcohol)+scale(BMI), data = DF3)
summary(model6)
screenreg(list(model6),
          custom.model.names = c("Alcohol and BMI"),
          custom.coef.names = c("Intercept", "Alcohol", "BMI"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))
```

### Question 6

Creating a model to see how schooling affects the life expectancy

```{r}
model7 <- lm(scale(Life_expectancy)~scale(Schooling), data = DF3)
summary(model7)
screenreg(list(model7),
          custom.model.names = c("Schooling"),
          custom.coef.names = c("Intercept", "Schooling"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))
```

Ploting of scatter plot with a regression line

```{r}
ggplot(DF2, aes(x = Schooling, y = Life_expectancy)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red")
```

### Question 7

Creating a model to see how alcohol affects the life expectancy

```{r}
model8 <- lm(scale(Life_expectancy)~scale(Alcohol), data = DF3)
summary(model8)
screenreg(list(model8),
          custom.model.names = c("Alcohol"),
          custom.coef.names = c("Intercept", "Alcohol"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))
```

### Question 8

Computing mean of life expectancy in densely populated countries

```{r}
mean(Dfk$Life_expectancy)
```

Computing mean of life expectancy in non-densely populated countries

```{r}
mean(DFa$Life_expectancy)
```

Additional data preparation part for visualization

```{r}
DFmean <- data.frame(Country=c("Densley Countries", "Non-densley Countries"),
                Life_Expectancy=c(76.28, 69.99))
head(df)

```

Creating bar graph

```{r}
ggplot(data=DFmean, aes(x=Country, y=Life_Expectancy)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Life_Expectancy), vjust=1.6, color="white", size=3.5)+
  theme_minimal()

```

### Question 9

Creating a model to see how immunization affects life expectancy

```{r}
model9 <- lm(scale(Life_expectancy)~scale(Hepatitis_B)+scale(Diphtheria)+scale(Polio), data = DF3)
summary(model9)
screenreg(list(model9),
          custom.model.names = c("Model 9"),
          custom.coef.names = c("Intercept", "Hepatitis_B","Diphtheria","Polio"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))
```

Creating three models (He,Di,Po) to see how immunization would help Life expectancy

```{r}
modelHe <- lm(scale(Life_expectancy)~scale(Hepatitis_B), data = DF3)
summary(modelHe)
screenreg(list(modelHe),
          custom.model.names = c("Hepatitis B"),
          custom.coef.names = c("Intercept", "Hepatitis_B"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))

modelDi <- lm(scale(Life_expectancy)~+scale(Diphtheria), data = DF3)
summary(modelDi)
screenreg(list(modelDi),
          custom.model.names = c("Diphtheria"),
          custom.coef.names = c("Intercept", "Diphtheria"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))

modelPo <- lm(scale(Life_expectancy)~+scale(Polio), data = DF3)
summary(modelPo)
screenreg(list(modelPo),
          custom.model.names = c("Polio"),
          custom.coef.names = c("Intercept", "Polio"),
          digits = 6,
          custom.gof.names = c("R-squared","Adjusted R-Squared","Number of Observations"))
```
